Inverse energy transfer in three-dimensional quantum vortex flows

### 2025-04-22 initial_submission.tex
Added initial submission file for comparisons

A chiral imbalance occurs if the mutual friction force is helical, 
\textit{i.e.} if the ratio $|f^+|^2/|f^-|^2 \neq 1$, with $|f^\pm|^2$ the total squared norm of the mutual friction components. 
In Fig.~\ref{fig:mutual-friction-decomp}, we show the temporal evolution of $|f^+|^2/|f^-|^2$, for both temperatures. 
%
\begin{figure}[h!]
    \centering
    \includegraphics*[width=0.48\textwidth]{fmfDecompFig.pdf}
    \caption{Temporal evolution of projected mutual friction force components $f^\pm$/ Inset: temporal evolution of total helical components.}
    \label{fig:mutual-friction-decomp}
\end{figure}
%
It is apparent
that during and after the reconnection, the mutual friction force is strongly chiral,
injecting more negative helicity than positive helicity. 
As a result, the ratio $\mathcal{H}^+/\mathcal{H}^-$ 
(reported in the inset of Fig.~\ref{fig:mutual-friction-decomp}) 
decreases significantly at reconnection and remains smaller than unity even at later times, 
indicating that the flow is persistently chiral. We conclude that the reconnection triggers indeed a chiral imbalance.
From Fig.~\ref{fig:mutual-friction-decomp} we determine the non-dimensional timescale $\tau = (t^*-t_R)E_R^{1/2}/L_R$
during which the mutual friction force is chiral as a result of reconnections: $\tau \approx 0.01$ and $\tau \approx 0.005$
for $T=1.9$K and $T=2.1$K, respectively, corresponding dimensionally to $\tau \approx 0.1$s for both temperatures. 
In superfluid turbulence, the timescale between two consecutive reconnections can be smaller than $\tau$ provided that the
vortex line density $\mathcal{L}$ (length of vortices per unit volume) is larger than $10^8 \text{m}^{-2}$ \cite{stasiak2024quantum,barenghi2004},
a condition which is easily met in superfluid helium experiments \cite{roche2007,Babuin2014}.

@@

\begin{thebibliography}{61}%

@@

      {\bibinfo  {journal} {Phys. Rev. Letters}\ }\textbf {\bibinfo {volume}

@@

      {\bibinfo  {journal} {Proc Natl Acad Sci Usa}\ }\textbf {\bibinfo {volume}

@@

    \bibitem [{\citenamefont {Schwarz}(1988)}]{schwarz1988}%
      \bibfield  {author} {\bibinfo {author} {\bibfnamefont {{\relax
      KW}.}~\bibnamefont {Schwarz}},\ }\bibfield  {title} {\bibinfo {title}
      {Three-dimensional vortex dynamics in superfluid {$^{4}$}{{He}}},\
      }\textbf {\bibinfo {volume} {38}},\ \bibinfo {pages} {2398} (\bibinfo {year}
      {1988})}\BibitemShut {NoStop}%
    \bibitem [{\citenamefont {Proment}\ and\ \citenamefont
      {Krstulovic}(2020)}]{promentMatchingTheoryCharacterize2020a}%
      \bibfield  {author} {\bibinfo {author} {\bibfnamefont {D.}~\bibnamefont
      {Proment}}\ and\ \bibinfo {author} {\bibfnamefont {G.}~\bibnamefont
      {Krstulovic}},\ }\bibfield  {title} {\bibinfo {title} {Matching theory to
      characterize sound emission during vortex reconnection in quantum fluids},\
      }\href {https://doi.org/10.1103/PhysRevFluids.5.104701} {\bibfield  {journal}
      {\bibinfo  {journal} {Phys. Rev. Fluids}\ }\textbf {\bibinfo {volume} {5}},\
      \bibinfo {pages} {104701} (\bibinfo {year} {2020})}\BibitemShut {NoStop}%
    \bibitem [{\citenamefont
      {Baggaley}(2012)}]{baggaleySensitivityVortexFilament2012a}%
      {Baggaley}},\ }\bibfield  {title} {\bibinfo {title} {The sensitivity of the
      vortex filament method to different reconnection models},\ }\href@noop {}
      {\bibfield  {journal} {\bibinfo  {journal} {J. Low Temp. Phys.}\ }\textbf
      {\bibinfo {volume} {168}},\ \bibinfo {pages} {18} (\bibinfo {year}
    \bibitem [{\citenamefont {Galantucci}\ \emph {et~al.}(2015)\citenamefont
      {Galantucci}, \citenamefont {Sciacca},\ and\ \citenamefont
      {Barenghi}}]{galantucciCoupledNormalFluid2015a}%
      {Galantucci}}, \bibinfo {author} {\bibfnamefont {M.}~\bibnamefont
      {Sciacca}},\ and\ \bibinfo {author} {\bibfnamefont {{\relax
      CF}.}~\bibnamefont {Barenghi}},\ }\bibfield  {title} {\bibinfo {title}
      {Coupled normal fluid and superfluid profiles of turbulent helium {{II}} in
      channels},\ }\href@noop {} {\bibfield  {journal} {\bibinfo  {journal} {Phys
      Rev B}\ }\textbf {\bibinfo {volume} {92}},\ \bibinfo {pages} {174530}
      (\bibinfo {year} {2015})}\BibitemShut {NoStop}%
    \bibitem [{\citenamefont {Gualtieri}\ \emph {et~al.}(2015)\citenamefont
      {Gualtieri}, \citenamefont {Picano}, \citenamefont {Sardina},\ and\
      \citenamefont {Casciola}}]{gualtieri2015exact}%
      {Gualtieri}}, \bibinfo {author} {\bibfnamefont {F.}~\bibnamefont {Picano}},
      \bibinfo {author} {\bibfnamefont {G.}~\bibnamefont {Sardina}},\ and\ \bibinfo
      {author} {\bibfnamefont {C.~M.}\ \bibnamefont {Casciola}},\ }\bibfield
      {title} {\bibinfo {title} {Exact regularized point particle method for
      multiphase flows in the two-way coupling regime},\ }\href@noop {} {\bibfield
      {journal} {\bibinfo  {journal} {J. Fluid Mech.}\ }\textbf {\bibinfo {volume}
      {773}},\ \bibinfo {pages} {520} (\bibinfo {year} {2015})}\BibitemShut
    \bibitem [{\citenamefont {Gualtieri}\ \emph {et~al.}(2017)\citenamefont
      {Gualtieri}, \citenamefont {Battista},\ and\ \citenamefont
      {Casciola}}]{gualtieri2017turbulence}%
      {Gualtieri}}, \bibinfo {author} {\bibfnamefont {F.}~\bibnamefont
      {Battista}},\ and\ \bibinfo {author} {\bibfnamefont {C.~M.}\ \bibnamefont
      {Casciola}},\ }\bibfield  {title} {\bibinfo {title} {Turbulence modulation in
      heavy-loaded suspensions of tiny particles},\ }\href@noop {} {\bibfield
      {journal} {\bibinfo  {journal} {Phys. Rev. Fluids}\ }\textbf {\bibinfo
      {volume} {2}},\ \bibinfo {pages} {034304} (\bibinfo {year}

@@

\appendix
\section*{Supplementary Materials}

Using Schwarz mesoscopic model \cite{schwarz1988}, vortex lines can be described as space curves $\s(\xi,t)$ of infinitesimal thickness, with a single quantum of circulation $\kappa=h/m_4=9.97\times10^{-8}\text{m}^2/\text{s}$, where $h$ is Planck's constant, $m_4=6.65\times10^{-27}\text{kg}$ is the mass of one helium atom, $\xi$ is the natural parameterisation, arclength, and $t$ is time. These conditions are a good approximation, since the vortex core radius of superfluid \textsuperscript{4}He($a_0=10^{-10}\text{m}$) is much smaller than any of the length scales of interest in turbulent flows. The equation of motion is
\begin{equation}
	\dot{\s}(\xi,t) = \v_s + \frac{\beta}{1+\beta}\left[\v_{ns}\cdot \s'\right]\s' + \beta\s'\times\v_{ns}+\beta'\s'\times\left[\s'\times \v_{ns}\right],
\end{equation}
where $\dot{\s}=\partial\s/\partial t$, $\s'=\partial\s/\partial \xi$ is the unit tangent vector, $\v_{ns}=\v_n - \v_s$, $\v_n$ and $\v_s$ are the normal fluid and superfluid velocities at $\s$ and $\beta$,$\beta'$ are temperature and Reynolds number dependent mutual fricition coefficients \cite{galantucciNewSelfconsistentApproach2020b}. The superfluid velocity $\v_s$ at a point $\x$ is determined by the Biot-Savart law
\begin{equation}
	\v_s(\x,t) = \frac{\kappa}{4\pi}\oint_{\mathcal{T}}\frac{\s'(\xi,t)\times\left[\x-\s(\xi,t)\right]}{|\x-\s(\xi,t)|}d\xi,
\end{equation}
where $\mathcal{T}$ represents the entire vortex configuration.
There is currently a lack of a well-defined theory of vortex reconnections in superfluid helium, like for the Gross-Pitaevskii equation \cite{villois2020irreversible,villoisUniversalNonuniversalAspects2017,promentMatchingTheoryCharacterize2020a}. An \emph{ad hoc} vortex reconnection algorithm is employed to resolve the collisions of vortex lines \cite{baggaleySensitivityVortexFilament2012a}.

\begin{figure}
	\centering
	\begin{subfigure}[b]{0.49\textwidth}
		\centering
		\includegraphics[width=0.8\textwidth]{schematic.png}
	\end{subfigure}
	\begin{subfigure}[b]{0.49\textwidth}
		\centering
		\includegraphics[width=0.9\textwidth]{schematic-2.png}
	\end{subfigure}
	\caption{Schematic diagram of the initial vortex configuration.}
	\label{fig:schematic}
\end{figure}

A \emph{two-way model} is crucial to understand the accurately interept the back-reaction effect of the normal fluid on the vortex line and vice-versa \cite{stasiak2024quantum}. We self-consistently evolve the normal fluid $\v_n$ with a modified Navier-Stokes equation
\begin{equation}
	\frac{\partial \v_n}{\partial t} + (\v_n\cdot\nabla)\v_n = -\nabla\frac{p}{\rho} + \nu_n\nabla^2\v_n + \frac{\mathbf{F}_{ns}}{\rho_n},
\end{equation}
\begin{equation}
	\mathbf{F}_{ns} = \oint_{\mathcal{T}}\mathbf{f}_{ns}\delta(\x-\x)d\xi, \quad \nabla\cdot\v_n=0,
\end{equation}
where $\rho=\rho_n + \rho_s$ is the total density, $\rho_n$ and $\rho_s$ are the normal fluid and superfluid densities, $p$ is the pressure, $\nu_n$ is the kinematic viscosity of the normal fluid and $\mathbf{f}_{ns}$ is the local friction per unit length \cite{galantucciCoupledNormalFluid2015a}
\begin{equation}
	\mathbf{f}_{ns} = -\mathcal{D}\s'\times\left[\s'\times(\dot{\s}-\v_n)\right]-\rho_n\kappa\s'\times(\v_n-\dot{\s}), 
	\label{eq:mutual-friction}
\end{equation}
where $\mathcal{D}$ is a coefficient dependent on the vortex Reynolds number and intrinsic properties of the normal fluid. The regularisation of the mutual fricition force onto the normal fluid grid is physically motivated by the strongly localised injection of vorticity during the momentum exchange of point-like particles and viscous flow in classical fluid dynamics \cite{gualtieri2015exact,gualtieri2017turbulence}. In short, the localised vorticity induced by the relative motion between the vortex lines and the normal fluid is diffused to discretisation of the grid spacing $\Delta x$ in a time interval $\epsilon_R$. In this way, the delta-forced fricition as defined in Eq.~\ref{eq:mutual-friction} is regularised by a Gaussian function, the fundamental solution of the diffusion equation. Further details of the method for classical fluids are contained in \cite{gualtieri2015exact,gualtieri2017turbulence} and for FOUCAULT in \cite{galantucciNewSelfconsistentApproach2020b}.

In this Letter, we report all results using dimensionless units, where the characteristic length scale is $\tilde{\lambda} = D/D_0$, where $D^3=(1\times10^{-3}\mathrm{m})^3$ is the dimensional cube size, $D_0^3=(2\pi)^3$ is the non-dimensional cubic computational domain. The time scale is given by $\tilde{\tau}=\tilde{\lambda}^2\nu_n^0\nu_n$, where the non-dimensional viscosity $\nu_n^0$ resolves the small scales of the normal fluid. In these simulations, these quanties are $\tilde{\lambda}=1.59\times10^{-4}$m, $\nu_n^0=0.32$ and $\tilde{\tau}=0.366$s at $T=1.9K$ and $\tilde{\tau}=0.485$s at $T=2.1K$. We consider an initial configuration of two pairs of orthongal vortices, initialised as shown in the schematic of Fig.~\ref{fig:schematic}. The seperation between vortices in each pair $d$ is set to be $d_v=0.5$ in dimensionless units, and the shortest distance between pairs is $D_v=\sqrt{(\pi-d_v/2)^2+\pi^2}\approx 3.08$, so that $d_v\ll D_v$. The Lagrangian discretisation of vortex lines is $\Delta \xi = 0.025$ (a total of 1340 discretisation points across the 4 vortex lines), using a timestep of $\Delta t_{VF} = 5.56\times10^{-6}$. For the normal fluid, a total of $N=256^3$ mesh point were used, with a timestep of $\Delta t_{NS} = 45\Delta t_{VF}$.

\end{document}
%
% ****** End of file apssamp.tex ******

>>>>

A chiral imbalance occurs if the mutual friction force is helical, 
\textit{i.e.} if the ratio $|f^+|^2/|f^-|^2 \neq 1$, with $|f^\pm|^2$ the total squared norm of the mutual friction components. 
In Fig.~\ref{fig:mutual-friction-decomp}, we show the temporal evolution of $|f^+|^2/|f^-|^2$, for both temperatures. 
%
\begin{figure}[h!]
    \centering
    \includegraphics*[width=0.48\textwidth]{fmfDecompFig.pdf}
    \caption{Temporal evolution of projected mutual friction force components $f^\pm$/ Inset: temporal evolution of total helical components.}.
    \label{fig:mutual-friction-decomp}
\end{figure}
%
It is apparent
that during and after the reconnection, the mutual friction force is strongly chiral,
injecting more negative helicity than positive helicity. 
As a result, the ratio $\mathcal{H}^+/\mathcal{H}^-$ 
(reported in the inset of Fig.~\ref{fig:mutual-friction-decomp}) 
decreases significantly at reconnection and remains smaller than unity even at later times, 
indicating that the flow is persistently chiral. We conclude that the reconnection triggers indeed a chiral imbalance.
From Fig.~\ref{fig:mutual-friction-decomp} we determine the non-dimensional timescale $\tau = (t^*-t_R)E_R^{1/2}/L_R$
during which the mutual friction force is chiral as a result of reconnections: $\tau \approx 0.01$ and $\tau \approx 0.005$
for $T=1.9$K and $T=2.1$K, respectively, corresponding dimensionally to $\tau \approx 0.1$s for both temperatures. 
In superfluid turbulence, the timescale between two consecutive reconnections can be smaller than $\tau$ provided that the
vortex line density $\mathcal{L}$ (length of vortices per unit volume) is larger than $10^8 \text{m}^{-2}$ \cite{stasiak2024quantum,barenghi2004},
a condition which is easily met in superfluid helium experiments \cite{roche2007,Babuin2014}.

@@

\begin{thebibliography}{55}%

@@

  {\bibinfo  {journal} {Phys. Rev. Lett.}\ }\textbf {\bibinfo {volume}

@@

  {\bibinfo  {journal} {Proc. Natl. Acad. Sci. Usa}\ }\textbf {\bibinfo {volume}

@@

@@

\end{document}
%
% ****** End of file apssamp.tex ******
